The \dealii Library, Version 9.7

### 2025-08-19 9.7/paper.tex

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Bruno Blais,
    Marc Fehling,
    Rene Gassm\"{o}ller,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Bruno Turcksin,
    David Wells,
  },
  pdftitle={The deal.II Library, Version 9.7, 2025},
}

@@

@@

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
\item MPGMRES is a variant of the
  widely used generalized minimal residual method (GMRES)
  \cite{Saad1986gmres}, distinguished by the support for multiple
  preconditioners (thus, \textit{MP}GMRES)  during the course of the solution process
  \cite{Greif2016gmres}. In our implementation in the
  \texttt{SolverMPGMRES} class, the use of $N$ preconditioners leads
  to the construction of multiple Krylov spaces via
  $N$-variate, non-commuting polynomials of the preconditioners and the system
  matrix applied to a residual vector. \texttt{SolverMPGMRES}
  supports two strategies: (i) a full variant that constructs all possible
  combinations of preconditioner application to the initial residual $r$,
  which for two preconditioners $P_1, P_2$ looks as
  \[
    r, P_1r, P_2r, P_1AP_1r, P_2AP_1r, P_1AP_2r, P_2AP_2r, P_1AP_1AP_1r,
   P_2AP_1AP_1r, \ldots, P_2AP_2AP_2r, \ldots;
 \]
 and (ii) a truncated version where no combinations of preconditioners are considered, i.e.
 \[
    r, P_1r, P_2r, P_1AP_1r, P_2AP_2r, P_1AP_1AP_1r,
   P_2AP_2AP_2r, \ldots.
 \]
 The new solver is beneficial when two different preconditioners can help to
 more quickly span the space in which the solution is expected.
\item A common operation in finite element codes is to ask whether
  a specific shape function belongs to a particular solution
  variable. In the past, this was implemented by using the
  \texttt{FiniteElement::system\_to\_component\_index()} function that
  returns, among other information, which vector component a shape
  function belongs to. One then had to test whether this is the vector
  component (or one of the vector components) that corresponds to a
  specific variable. The new
  \texttt{FiniteElement::shape\_function\_belongs\_to()} makes this
  easier: It takes as argument a \texttt{FiniteElementExtractor} that
  corresponds to a scalar, vector-valued, or tensor-valued set of
  solution variables, and directly returns \texttt{true} or
  \texttt{false}.
\item Most of the functions in the \texttt{GridGenerator} namespace correspond
  to simple geometries (such as cubes, spheres, and other shapes with analytical
  formulas) discretized with hypercubes. The function
  \texttt{GridGenerator::convert\_hypercube\_to\_simplex\_mesh()} converts such
  discretizations to instead use simplices. That function now supports
  anisotropic splits, including splitting quadrilaterals into two triangles and
  hexahedra into six tetrahedra, in addition to the standard isotropic splits.
  These anisotropic splits significantly reduce the total number of cells and
  are useful in contexts where one would like to use as few cells as
  possible.
\item The build system now supports interprocedural and link-time optimization
  (LTO). This can be enabled with the CMake configuration option
  \texttt{-DDEAL\_II\_WITH\_LTO=ON}. This option enables whole-program
  optimization at link-time and can make executables significantly faster.
\item \texttt{GridIn::read\_vtk()} now loads cell-centered data vectors stored
  in the input file and makes them available via
  \texttt{GridIn::get\_cell\_data()}.
\item Sparsity patterns for matrices built with \texttt{FE\_Q\_iso\_Q1} elements
  now correctly reflect the inherited sparsity on each cell, making the usage of
  this element usable as a preconditioner for a higher order problems (this is known
  as low order rediscretization, see also \cite{pazner2023low}).
\end{itemize}
%
The
\href{https://dealii.org/current/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}{changelog}
-- listing more than 120
features and bugfixes --
contains a complete record of all changes; see \cite{changes97}.

@@

\begin{quote}
%  authors removed:
%  Daniel       Arndt,
%  Wolfgang     Bangerth,
%  Bruno        Blais,
%  Marc         Fehling,
%  Rene         Gassmoeller,
%  Timo         Heister,
%  Luca         Heltai,
%  Martin       Kronbichler,
%  Matthias     Maier,
%  Peter        Munch,
%  Jean-Paul    Pelteret,
%  Bruno        Turcksin,
%  David        Wells,
Pasquale     Africa,
Henry        Arhin,
Maximilian   Bergbauer,
Alfredo      Buttari,
Bruna        Campos,
Nicholas     Cantrell,
Xiaoming     Cao,
Chayapol     Chaoveeraprasit,
Jerett       Cherry,
Dario        Coscia,
John         Coughlin,
Nikita       Daniliuk,
Crystal      Farris,
Marco        Feder,
Emmanuel     Ferdman,
Federico     Fernandez,
Olivier      Gaboriault,
Mohamad      Ghadban,
Mahdi        Gharehbaygloo,
Robin        Görmer,
Davit        Gyulamiryan,
Lóránt       Hadnagy,
Fernando     Herrera,
Robin        Hiniborch,
Quang        Hoang,
Jordan       Hoffart,
Sascha       Hofstetter,
Yimin        Jin,
Yann         Jobic,
Sean         Johnson,
Vaishnavi    Kale,
Sebastian    Kinnewig,
Andreas      Koch,
Jason        Landini,
Zhou         Lei,
Yingli       Li,
Nils         Margenberg,
Oreste       Marquis,
Ryan         Moulday,
Nils         Much,
Tileuzhan    Mukhamet,
Wasim~Niyaz  Munshi,
Natalia      Nebulishvili,
Luz          Paz,
David        Pecoraro,
Davide       Polverino,
Sanjeeb      Poudel,
Guilhem      Poy,
Laura        Prieto~Saavedra,
Sebastian    Proell,
Andreas      Ritthaler,
Mayank       Sabharwal,
Sam          Scheuerman,
Magdalena    Schreter,
Richard      Schussnig,
Kyle         Schwiebert,
Marc         Secanell,
Qingyuan     Shi,
Wyatt        Smith,
Simon        Sticko,
Dominik      Still,
Edward       Terrell,
Jan~Philipp  Thiele,
Xiaochuan    Tian,
Siarhei      Uzunbajakau,
Mikael       Vaillant,
Vinayak      Vijay,
Stephan      Voss,
Michał       Wichrowski,
Simon        Wiesheier,
Yi-Yung      Yang,
\end{quote}
Their contributions are much appreciated -- this project lives by its
community of users and contributors!

>>>>

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Bruno Blais,
    Marc Fehling,
    Rene Gassm\"{o}ller,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Bruno Turcksin,
    David Wells,
    Michał Wichrowski,
  },
  pdftitle={The deal.II Library, Version 9.7, 2025},
}

@@

  \author[13]{Michał Wichrowski}
\affil[13]{Interdisciplinary Center for Scientific Computing, Heidelberg University, Heidelberg, Germany.
  {\texttt{mt.wichrowsk@uw.edu.pl}}}

@@

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
\item MPGMRES is a variant of the
  widely used generalized minimal residual method (GMRES)
  \cite{Saad1986gmres}, distinguished by the support for multiple
  preconditioners (thus, \textit{MP}GMRES)  during the course of the solution process
  \cite{Greif2016gmres}. In our implementation in the
  \texttt{SolverMPGMRES} class, the use of $N$ preconditioners leads
  to the construction of multiple Krylov spaces via
  $N$-variate, non-commuting polynomials of the preconditioners and the system
  matrix applied to a residual vector. \texttt{SolverMPGMRES}
  supports two strategies: (i) a full variant that constructs all possible
  combinations of preconditioner application to the initial residual $r$,
  which for two preconditioners $P_1, P_2$ looks as
  \[
    r, P_1r, P_2r, P_1AP_1r, P_2AP_1r, P_1AP_2r, P_2AP_2r, P_1AP_1AP_1r,
   P_2AP_1AP_1r, \ldots, P_2AP_2AP_2r, \ldots;
 \]
 and (ii) a truncated version where no combinations of preconditioners are considered, i.e.
 \[
    r, P_1r, P_2r, P_1AP_1r, P_2AP_2r, P_1AP_1AP_1r,
   P_2AP_2AP_2r, \ldots.
 \]
 The new solver is beneficial when two different preconditioners can help to
 more quickly span the space in which the solution is expected.
\item A common operation in finite element codes is to ask whether
  a specific shape function belongs to a particular solution
  variable. In the past, this was implemented by using the
  \texttt{FiniteElement::system\_to\_component\_index()} function that
  returns, among other information, which vector component a shape
  function belongs to. One then had to test whether this is the vector
  component (or one of the vector components) that corresponds to a
  specific variable. The new
  \texttt{FiniteElement::shape\_function\_belongs\_to()} makes this
  easier: It takes as argument a \texttt{FiniteElementExtractor} that
  corresponds to a scalar, vector-valued, or tensor-valued set of
  solution variables, and directly returns \texttt{true} or
  \texttt{false}.
\item Most of the functions in the \texttt{GridGenerator} namespace correspond
  to simple geometries (such as cubes, spheres, and other shapes with analytical
  formulas) discretized with hypercubes. The function
  \texttt{GridGenerator::convert\_hypercube\_to\_simplex\_mesh()} converts such
  discretizations to instead use simplices. That function now supports
  anisotropic splits, including splitting quadrilaterals into two triangles and
  hexahedra into six tetrahedra, in addition to the standard isotropic splits.
  These anisotropic splits significantly reduce the total number of cells and
  are useful in contexts where one would like to use as few cells as
  possible.
\item The build system now supports interprocedural and link-time optimization
  (LTO). This can be enabled with the CMake configuration option
  \texttt{-DDEAL\_II\_WITH\_LTO=ON}. This option enables whole-program
  optimization at link-time and can make executables significantly faster.
\item \texttt{GridIn::read\_vtk()} now loads cell-centered data vectors stored
  in the input file and makes them available via
  \texttt{GridIn::get\_cell\_data()}.
\item Sparsity patterns for matrices built with \texttt{FE\_Q\_iso\_Q1} elements
  now correctly reflect the inherited sparsity on each cell, making the usage of
  this element usable as a preconditioner for a higher order problems (this is known
  as low order rediscretization, see also \cite{pazner2023low}).
  \item The assembly of ghost penalty stabilization matrices, which appear in CutFEM, can now be done 
    via \texttt{TensorProductMatrixCreator::create\_1d\_ghost\_penalty\_matrix}. For Cartesian cells, the local penalty matrix can be constructed by exploiting the tensor-product structure   
    of the basis functions. The ghost penalty operator is expressed as 
    a Kronecker product of precomputed one-dimensional mass and penalty matrices,
    which avoids the evaluation of high-order derivatives~\cite{wichrowski2025matrix}.
    The new functionality also handles the necessary lexicographical renumbering 
    of degrees of freedom across adjacent cells via \texttt{FETools::cell\_to\_face\_patch}.
\end{itemize}
%
The
\href{https://dealii.org/current/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}{changelog}
-- listing more than 120
features and bugfixes --
contains a complete record of all changes; see \cite{changes97}.

@@

\begin{quote}
%  authors removed:
%  Daniel       Arndt,
%  Wolfgang     Bangerth,
%  Bruno        Blais,
%  Marc         Fehling,
%  Rene         Gassmoeller,
%  Timo         Heister,
%  Luca         Heltai,
%  Martin       Kronbichler,
%  Matthias     Maier,
%  Peter        Munch,
%  Jean-Paul    Pelteret,
%  Bruno        Turcksin,
%  David        Wells,
Pasquale     Africa,
Henry        Arhin,
Maximilian   Bergbauer,
Alfredo      Buttari,
Bruna        Campos,
Nicholas     Cantrell,
Xiaoming     Cao,
Chayapol     Chaoveeraprasit,
Jerett       Cherry,
Dario        Coscia,
John         Coughlin,
Nikita       Daniliuk,
Crystal      Farris,
Marco        Feder,
Emmanuel     Ferdman,
Federico     Fernandez,
Olivier      Gaboriault,
Mohamad      Ghadban,
Mahdi        Gharehbaygloo,
Robin        Görmer,
Davit        Gyulamiryan,
Lóránt       Hadnagy,
Fernando     Herrera,
Robin        Hiniborch,
Quang        Hoang,
Jordan       Hoffart,
Sascha       Hofstetter,
Yimin        Jin,
Yann         Jobic,
Sean         Johnson,
Vaishnavi    Kale,
Sebastian    Kinnewig,
Andreas      Koch,
Jason        Landini,
Zhou         Lei,
Yingli       Li,
Nils         Margenberg,
Oreste       Marquis,
Ryan         Moulday,
Nils         Much,
Tileuzhan    Mukhamet,
Wasim~Niyaz  Munshi,
Natalia      Nebulishvili,
Luz          Paz,
David        Pecoraro,
Davide       Polverino,
Sanjeeb      Poudel,
Guilhem      Poy,
Laura        Prieto~Saavedra,
Sebastian    Proell,
Andreas      Ritthaler,
Mayank       Sabharwal,
Sam          Scheuerman,
Magdalena    Schreter,
Richard      Schussnig,
Kyle         Schwiebert,
Marc         Secanell,
Qingyuan     Shi,
Wyatt        Smith,
Simon        Sticko,
Dominik      Still,
Edward       Terrell,
Jan~Philipp  Thiele,
Xiaochuan    Tian,
Siarhei      Uzunbajakau,
Mikael       Vaillant,
Vinayak      Vijay,
Stephan      Voss,
% Michał       Wichrowski,
Simon        Wiesheier,
Yi-Yung      Yang,
\end{quote}
Their contributions are much appreciated -- this project lives by its
community of users and contributors!
